---
title: "Second Write Up on RUM model"
author: "Andie Creel"
date: "2023-07-06"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message  = FALSE)
```

# Questions of interest: 

  - Substitution between indoor and outdoor leisure under different climate scenarios 
  - How does that substitution change during seasons
  - How does it change for people living in an urban heat island 
  
Note: all results are on the extensive margin (did you take a trip that week or not). 

From the 05.write_up.pdf I found that constructing the travel cost for no trips is fairly robust, regardless of if I'm using total travel time or average travel time per trip, as well as if I'm grouping by race or state. 

In this write up, I use **total travel time grouped by state** to get the cost when people don't take a trip. 


# RUM model 

Indirect utility:

$$V_{itc} = \alpha_c + \beta x_{ic} + \gamma_c w_t $$
Where $V$ is the indirect utility for person $i$ on choice occasion $t$ (day) for choice $c$. $\alpha$ is a choice specific constant. $\beta$ is a general coefficient for travel **time** $x$. $\gamma$ is a choice specific coefficient for weather on day $t$ at for the choice occasion $w$. 


```{r}
rm(list = ls())
options(scipen = 999)
library(dplyr)
library(tidyr)
library(stringr)
library(vroom)
library(mlogit)
library(stargazer)
library(lubridate)
library(ggplot2)

# -----------------------------------------------------------------------------
# read in R memory built in R_data_objs
# -----------------------------------------------------------------------------
load("05.RUM_data_objs.RData")

# -----------------------------------------------------------------------------
# preferred reg specification (total travel time grouped by state)
# -----------------------------------------------------------------------------

# income not included
reg1.a <- mlogit(choice ~ travel_time_total_state |  tmmx, # formula
               myRUM_idx, #mlogit data object
               reflevel = "indoor_home", #reference level 
               ) # choices available (dropped no leisure)

# income included 
reg1.b <- mlogit(choice ~ travel_time_total_state | fam_inc_mid + tmmx, # formula
               myRUM_idx, 
               reflevel = "indoor_home") 

stargazer(reg1.a, reg1.b,
          title = "preferred model (entire year)",
          type = "text")

```
I'm currently going to work with the regression where income is not included because the estimates are robust. The numeraire is currently minutes. 

# Marginal Effects 

In multinomial logit models, the coefficients are NOT the marginal effects. 

Marginal effect equations: 
$$\frac{\partial P_{itl}}{\partial x_{itl}} = \beta P_{itl}(1 - P_{itl})$$

$$\frac{\partial P_{itl}}{\partial w_{t}} = P_{itl}(\gamma_l - \sum_c P_{itc} \gamma_c)$$

$$\frac{\partial P_{itl}}{\partial x_{itk}} = - \beta P_{itl}P_{itk}$$

## Elasticities for temperature 
```{r}
# -----------------------------------------------------------------------------
# preferred reg specification (total travel time grouped by state)
# -----------------------------------------------------------------------------
elasts.temp.1a <- effects(reg1.a, covariate = "tmmx", type = "rr")


```

A 1% increase in temperature leads: 

  - `r round(elasts.temp.1a["indoor_home"],2)`% change in probability of choosing indoor leisure at home 
  - `r round(elasts.temp.1a["indoor_away"],2)`% change in probability of choosing indoor leisure away from home 
  - `r round(elasts.temp.1a["outdoor_away"],2)`% change in probability of choosing outdoor leisure away from home
  - `r round(elasts.temp.1a["outdoor_home"],2)`% change in probability of choosing outdoor leisure at home 
  
Note: I find it reassuring that the outdoor leisure at home and away from home increase in a similar way. 

## Elasticities for travel time

```{r}
elasts.travel.1a<- effects(reg1.a, covariate = "travel_time_total_state", type = "rr")
elasts.travel.1a
```

A 1% increase in travel time to indoor away from home activities leads to: 

  - `r round(elasts.travel.1a["indoor_away", "indoor_home"], 2)`% change in probability of choosing indoor home, outdoor away and outdoor home leisure (IIA property)
  - `r round(elasts.travel.1a["indoor_away", "indoor_away"], 2)`% change in probability of choosing indoor away from leisure 
  
  
A 1% increase in travel time to outdoor away from home leisure activities leads to: 

  - `r round(elasts.travel.1a["outdoor_away", "indoor_home"], 3)`% change in probability of choosing indoor home, indoor away and outdoor home leisure (IIA property)
  - `r round(elasts.travel.1a["outdoor_away", "outdoor_away"], 2)`% change in probability of choosing outdoor away from home leisure 


# By Season 

```{r}
# -----------------------------------------------------------------------------
# Season regressions
# -----------------------------------------------------------------------------

mySeasonReg <- function(i){

  myWorking_temp <- myRUM_idx %>%
    filter(season == i) 
  
  reg <- mlogit(choice ~ travel_time_avg_state | tmmx, # formula
                         myWorking_temp, #mlogit data object
                         reflevel = "indoor_home") #reference level
  
}

seasonRegs <- lapply(c("spring","summer","fall","winter"), 
                      mySeasonReg)

stargazer(seasonRegs, 
          column.labels = c("spring","summer","fall","winter"),
          title = "The effect of temperature changes given the season.",
          type = "text")



```

## marginal effects 

### Temperature 

```{r}
# -----------------------------------------------------------------------------
# seasonal marginal effects
# -----------------------------------------------------------------------------

myEffects <- function(x){
  effects(x, covariate = "tmmx", type = "rr")
}

season_temp_elasts <- lapply(seasonRegs, myEffects) %>%
  bind_rows() %>%
  round(2)

season_temp_elasts <- bind_cols(season = c("spring","summer","fall","winter"), 
                                season_temp_elasts)


```


# Have these summer outdoor_away elasticities trended through time? 


```{r}
# -----------------------------------------------------------------------------
# getting summer elasticity through time
# -----------------------------------------------------------------------------

getSummerElasticity <- function(y) {
  
  myWorking_temp <- myRUM_idx %>%
    filter(season == "summer") %>%
    filter(year == y)
  
  reg <- mlogit(choice ~ travel_time_avg_state | tmmx, # formula
                         myWorking_temp, #mlogit data object
                         reflevel = "indoor_home") #reference level
  
  effects(reg, covariate = "tmmx", type = "rr")
}

#run regs and turn into table
season_temp_elasts_time <- lapply(2003:2021, getSummerElasticity) %>% 
  bind_rows()

# add in year
season_temp_elasts_time <- bind_cols(year = 2003:2021, 
                                      season_temp_elasts_time)

# -----------------------------------------------------------------------------
# graph
# -----------------------------------------------------------------------------

seasons_elasts_long <- season_temp_elasts_time %>%
  pivot_longer(
    cols = c(outdoor_away,outdoor_home, indoor_away,  indoor_home), # Select the columns to pivot
    names_to = "activity",      
    values_to = "temp_elasticity"          
  )


# Plotting the data using ggplot
 ggplot(seasons_elasts_long, aes(x = year, y = temp_elasticity, color = activity)) +
  geom_point() +
  geom_smooth(method = "lm", se = F) +
  labs(x = "Year", y = "Elasticity (temperature increase)") +
  scale_color_discrete(name = "activity") +
  ggtitle("Summer elasticities to temperature increases") +
  theme_bw()


```
An increase summer temperature drives people home. This trend has only been exaggerated through time (potentially as the world has warmed through the last 20 years).
