---
title: "RUM Model: welfare calcs"
author: "Andie Creel"
date: "2023-07-06"
output: 
  pdf_document:
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message  = FALSE)
```

# Questions of interest: 

As summers heat up, how can individuals adapt the time of year they recreate outside? How can institutions like schools and work change so that people can adapt their recreation schedule. 
  
# RUM model (Annual)

Indirect utility:

$$V_{itc} = \alpha_c + \beta x_{ic} + \gamma_c w_t $$
Where $V$ is the indirect utility for person $i$ on choice occasion $t$ (day) for choice $c$. $\alpha$ is a choice specific constant. $\beta$ is a general coefficient for travel **time** $x$. $\gamma$ is a choice specific coefficient for weather on day $t$ at for the choice occasion $w$. 


```{r}
rm(list = ls())
options(scipen = 999)
library(dplyr)
library(tidyr)
library(stringr)
library(vroom)
library(mlogit)
library(stargazer)
library(lubridate)
library(ggplot2)
library(gridExtra)

# -----------------------------------------------------------------------------
# read in R memory built in R_data_objs
# -----------------------------------------------------------------------------
load("05.RUM_data_objs.RData")

# -----------------------------------------------------------------------------
# preferred reg specification (total travel time grouped by state)
# -----------------------------------------------------------------------------

# income not included
reg1 <- mlogit(choice ~ travel_time_total_state |  tmmx, 
               myRUM_idx, 
               reflevel = "indoor_home") 

stargazer(reg1,
          title = "preferred model (entire year)",
          type = "text")

```
 The numeraire is minutes. 

## Welfare calcs

We can identify consumer surplus up to an unidentified constant, therefore the number on its own is not relevant. However changes in consumer surplus are identified. 

```{r}
# -----------------------------------------------------------------------------
# calculate inclusive value for baseline
# -----------------------------------------------------------------------------
iv_baseline <- logsum(reg1)

# -----------------------------------------------------------------------------
# create a data set where travel time isn't ever more than 10 minutes for outdoor leisure
# -----------------------------------------------------------------------------

myRUM_10min <- myRUM_df %>%
  select(choice, travel_time_total_state, tmmx, activity) %>% 
  mutate(travel_time_total_state = if_else(activity == "outdoor_away" & travel_time_total_state > 10, 
                                           10, travel_time_total_state)) %>% 
  dfidx(idx = list(NA, "activity")) 

# calc indirect utility for everyone under new scenario 
iv_10min <- logsum(reg1, data = myRUM_10min)

# -----------------------------------------------------------------------------
# calculate change in welfare
# -----------------------------------------------------------------------------
surplus <- - (iv_10min - iv_baseline) / coef(reg1)["travel_time_total_state"] 
summary(surplus)



```


Consumer's surplus variation range from 0 to 19 minutes, with a median value of about 1 minute. This is per choice occasion (day). 








